Supporting Information:\\ NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2018-08-17 Manuscript/manuscriptPS.tex
Update text and figures

   % The salt concentrations calculated as [salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M.

@@

    \todo{The used definition of ion concentrations needs to be specified and standardized.} 

@@

  lipid/counter-ions & force field for lipids / ions & \footnote{Excess Na$^+$ or K$^+$ concentration}C$_{\rm ci}$ (M) & CaCl$_2$\,(M)  &  \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\

@@

    POPC:POPS (5:1)        & CHARMM36 \cite{klauda10,venable13,kim16}  &0 & 0.15 \todoi{Concentration to be checked} & 250:50 & ?  & ?  & 298  & 200 & ?  & \cite{??} \todoi{Trajectories and further details to be added by J. Madsen}  \\
    POPC:POPS (5:1)        & CHARMM36 \cite{klauda10,venable13,kim16}  &0 & 1 \todoi{Concentration to be checked} & 250:50 & ?  & ?  & 298  & 200 & ?  & \cite{??} \todoi{Trajectories and further details to be added by J. Madsen}  \\

@@

    POPC:OPPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &0    & 0.1 & 120:24 & 5760 & 10   & 298  & 200 & 200 \todoi{Equilibration?} & \cite{POPCpopsMACROG}  \\
    POPC:OPPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &0    & 0.3 & 120:24 & 5760 & 31   & 298  & 200 & 200 \todoi{Equilibration?} & \cite{POPCpopsMACROG}  \\
    POPC:OPPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &0    & 1   & 120:24 & 5760 & 104  & 298  & 200 & 200 \todoi{Equilibration?} & \cite{POPCpopsMACROG}  \\
    POPC:OPPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &0    & 3   & 120:24 & 5760 & 311  & 298  & 200 & 200 \todoi{Equilibration?} & \cite{POPCpopsMACROG}  \\
    POPC:OPPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &0.5  & 0   & 120:24 & 5760 & 52   & 298  & 200 & 190 & \cite{POPCpopsMACROGwithK}  \\
    POPC:OPPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &1    & 0   & 120:24 & 5760 & 104  & 298  & 200 & 190 & \cite{POPCpopsMACROGwithK}  \\
    POPC:OPPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &2    & 0   & 120:24 & 5760 & 208  & 298  & 200 & 145 & \cite{POPCpopsMACROGwithK}  \\
    POPC:OPPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &3    & 0   & 120:24 & 5760 & 311  & 298  & 200 & 125 & \cite{POPCpopsMACROGwithK}  \\
    POPC:OPPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &4    & 0   & 120:24 & 5760 & 415  & 298  & 200 & 125 & \cite{POPCpopsMACROGwithK}  \\

@@

    POPC:POPS (5:1)/K$^+$  & Lipid14/17 \cite{dickson14,gould18} &0.5  & 0   & 120:24 & 5760 & ?   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withK}  \\
    POPC:POPS (5:1)/K$^+$  & Lipid14/17 \cite{dickson14,gould18} &1    & 0   & 120:24 & 5760 & ?   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withK}  \\
    POPC:POPS (5:1)/K$^+$  & Lipid14/17 \cite{dickson14,gould18} &2    & 0   & 120:24 & 5760 & ?   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withK}  \\
    POPC:POPS (5:1)/K$^+$  & Lipid14/17 \cite{dickson14,gould18} &3    & 0   & 120:24 & 5760 & ?   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withK}  \\
    POPC:POPS (5:1)/K$^+$  & Lipid14/17 \cite{dickson14,gould18} &4    & 0   & 120:24 & 5760 & ?   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withK}  \\

@@

    POPC:POPS (5:1)/Na$^+$  & Lipid14/17 \cite{dickson14,gould18} &0.5  & 0   & 120:24 & 5760 & ?   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withNa}  \\
    POPC:POPS (5:1)/Na$^+$  & Lipid14/17 \cite{dickson14,gould18} &1    & 0   & 120:24 & 5760 & ?   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withNa}  \\
    POPC:POPS (5:1)/Na$^+$  & Lipid14/17 \cite{dickson14,gould18} &2    & 0   & 120:24 & 5760 & ?   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withNa}  \\
    POPC:POPS (5:1)/Na$^+$  & Lipid14/17 \cite{dickson14,gould18} &3    & 0   & 120:24 & 5760 & ?   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withNa}  \\
    POPC:POPS (5:1)/Na$^+$  & Lipid14/17 \cite{dickson14,gould18} &4    & 0   & 120:24 & 5760 & ?   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withNa}  \\

@@

    POPC:POPS (5:1)/Na$^+$ & Lipid14/17 \cite{dickson14,gould18} &0    & 0.03 & 60:12 & ?     & 0   & 298  & ?  & ?  & \cite{??} \todoi{Data to be delivered by Melcr}  \\
    POPC:POPS (5:1)/Na$^+$ & Lipid14/17 \cite{dickson14,gould18} &0    & 0.17 & 60:12 & ?     & 0   & 298  & ?  & ?  & \cite{??} \todoi{Data to be delivered by Melcr}  \\
    POPC:POPS (5:1)/Na$^+$ & Lipid14/17 \cite{dickson14,gould18} &0    & 0.36 & 60:12 & ?     & 0   & 298  & ?  & ?  & \cite{??} \todoi{Data to be delivered by Melcr}  \\

@@

    POPC:POPS (4:1)/Na$^+$  & Berger \cite{tieleman99,mukhopadhyay04}\todoi{Are these correct references?} &1    & 0   & 102:26 & 4290 & 80  & 310  & 200 & 50 & \cite{POPCpopsBERGERwith1000mMNa}  \\
    POPC:POPS (4:1)  & Berger \cite{tieleman99,mukhopadhyay04} &  0  & 0.12   & 104:24 & 4306 & 24 & 310  & 300 & 100 & \cite{POPCpopsBERGERwith102mMCa}  \\
    POPC:POPS (4:1)  & Berger \cite{tieleman99,mukhopadhyay04} &  0  & 0.715  & 104:24 & 4306 & 72 & 310  & 300 & 100 & \cite{POPCpopsBERGERwith715mMCa}  \\

@@

\todo{How close are the hydration levels in experiments and simulations?} \\
\todo{We need to decide how to report the ion concentrations, see discussion in Ref.~\citenum{melcr18}.}

@@

To evaluate counterion binding in different simulation models against experimental data~\cite{roux90},
we plot the headgroup order parameters measured from POPC:POPS 5:1 mixture
as a function of different monovalent ions added to the buffer (Fig. \ref{PSresponseTONaCl}). 
Experimental order parameter data for POPC headgroup in the mixture is available as a function
of LiCl and KCl concentrations, while POPS headgroup order parameters are measured also
as a function of NaCl. Lithium interacts more strongly with PS headgroups than other monovalent 
ions~\cite{hauser83,hauser85,roux86,mattai89,roux90}, as also observed for PC headgroups~\cite{cevc90}. 
This is evident also in the changes of PS headgroup order parameters, which decrease with the addition of lithium 
but increase with the addition of sodium or potassium (Fig. \ref{PSresponseTONaCl}). 
POPC headgroup order parameters exhibit a clear decrease as a function of LiCl concentration
but only modest changes as a function of KCl concentration, indicating singificant 
Li$^+$ binding but only weak Na$^+$ binding to the mixture when interpreted using the
electrometer concept~\cite{akutsu81,altenbach84,seelig87}. In simulations with the
Berger model, the headgroup order parameter response of POPC
to the added NaCl is similar to the experiments of LiCl,
indicating overestimated binding affinity of sodium, in line with the results for PC bilayers \cite{catte16}.
Indeed, the sodium density profile shows a significant binding peak in the 
Berger model (Fig. \ref{CIdensPSOCmixt}). Potassium binding in the MacRog simulation
is significantly weaker  (Fig. \ref{CIdensPSOCmixt}) and the headgroup order parameter 
changes are also in better agreement with simulations (Fig. \ref{PSresponseTONaCl}).
\todo{Discussion about Lipid17 to be written when we have the density profiles.}
All the tested models overestimate the changes of POPS headgroup order parameters as
a function of monovalent ions (Fig. \ref{PSresponseTONaCl}), suggesting that
model development is necessary to interpret the PS headgroup-ion interactions 
from MD simulations.
\begin{figure*}[]
  \centering
  \includegraphics[width=18.0cm]{../Figs/CHANGESwithMONVALENTwithPS.eps}
  \caption{\label{PSresponseTONaCl}
    Changes of the PC (left) and PS (right) headgroup order parameters as a function of
    added NaCl, KCl and LiCl from POPC:POPS (5:1) mixture. The experimental data is from Ref. \citenum{roux90}.
    The values from counterion-only systems are set as a zero point of y-axis.
    To correctly illustrate the significant forking of the $\alpha$-carbon order parameter
    in PS headgroup (bottom, right), the y-axis is tranferred with the same value for both order parameters such that the lower order
    parameter value is at zero.
  }
  \todo{CHARMM36 results for this plot would be highly useful.}
\end{figure*}

>>>>

  The salt concentrations calculated as [salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M.
  these correspond the concentrations reported in the experiments by Roux et al.~\cite{roux90}.

@@

@@

  lipid/counter-ions & force field for lipids / ions & \footnote{Excess Na$^+$ or K$^+$ concentration}C$_{\rm ci}$ (M) & [CaCl$_2$]\,(M)  &  \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\

@@

    POPC:POPS (5:1)        & CHARMM36 \cite{klauda10,venable13,kim16}  &0 & 0.15 \todoi{Concentration to be checked after the amount of water molecules is known.} & 250:50 & ?  & ?  & 298  & 200 & ?  & \cite{??} \todoi{Trajectories and further details to be added by J. Madsen}  \\
    POPC:POPS (5:1)        & CHARMM36 \cite{klauda10,venable13,kim16}  &0 & 1 \todoi{Concentration to be checked after the amount of water molecules is known.} & 250:50 & ?  & ?  & 298  & 200 & ?  & \cite{??} \todoi{Trajectories and further details to be added by J. Madsen}  \\

@@

    POPC:OPPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &0    & 0.10 & 120:24 & 5760 & 10   & 298  & 200 & 200 \todoi{Equilibration?} & \cite{POPCpopsMACROG}  \\
    POPC:OPPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &0    & 0.30 & 120:24 & 5760 & 31   & 298  & 200 & 200 \todoi{Equilibration?} & \cite{POPCpopsMACROG}  \\
    POPC:OPPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &0    & 1.00   & 120:24 & 5760 & 104  & 298  & 200 & 200 \todoi{Equilibration?} & \cite{POPCpopsMACROG}  \\
    POPC:OPPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &0    & 3.00   & 120:24 & 5760 & 311  & 298  & 200 & 200 \todoi{Equilibration?} & \cite{POPCpopsMACROG}  \\
    POPC:OPPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &0.50  & 0   & 120:24 & 5760 & 52   & 298  & 200 & 190 & \cite{POPCpopsMACROGwithK}  \\
    POPC:OPPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &1.00    & 0   & 120:24 & 5760 & 104  & 298  & 200 & 190 & \cite{POPCpopsMACROGwithK}  \\
    POPC:OPPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &2.00    & 0   & 120:24 & 5760 & 208  & 298  & 200 & 145 & \cite{POPCpopsMACROGwithK}  \\
    POPC:OPPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &3.00    & 0   & 120:24 & 5760 & 311  & 298  & 200 & 125 & \cite{POPCpopsMACROGwithK}  \\
    POPC:OPPS (5:1)/K$^+$  & MacRog \cite{maciejewski14} &4.00    & 0   & 120:24 & 5760 & 415  & 298  & 200 & 125 & \cite{POPCpopsMACROGwithK}  \\

@@

    POPC:POPS (5:1)/K$^+$  & Lipid14/17 \cite{dickson14,gould18} &0.5\todoi{Concentration to be checked after the amount of ions is known.}  & 0   & 120:24 & 5760 & ?   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withK}  \\
    POPC:POPS (5:1)/K$^+$  & Lipid14/17 \cite{dickson14,gould18} &1\todoi{Concentration to be checked after the amount of ions is known.}    & 0   & 120:24 & 5760 & ?   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withK}  \\
    POPC:POPS (5:1)/K$^+$  & Lipid14/17 \cite{dickson14,gould18} &2\todoi{Concentration to be checked after the amount of ions is known.}    & 0   & 120:24 & 5760 & ?   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withK}  \\
    POPC:POPS (5:1)/K$^+$  & Lipid14/17 \cite{dickson14,gould18} &3\todoi{Concentration to be checked after the amount of ions is known.}    & 0   & 120:24 & 5760 & ?   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withK}  \\
    POPC:POPS (5:1)/K$^+$  & Lipid14/17 \cite{dickson14,gould18} &4\todoi{Concentration to be checked after the amount of ions is known.}    & 0   & 120:24 & 5760 & ?   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withK}  \\

@@

    POPC:POPS (5:1)/Na$^+$  & Lipid14/17 \cite{dickson14,gould18} &0.5\todoi{Concentration to be checked after the amount of ions is known.}  & 0   & 120:24 & 5760 & ?   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withNa}  \\
    POPC:POPS (5:1)/Na$^+$  & Lipid14/17 \cite{dickson14,gould18} &1\todoi{Concentration to be checked after the amount of ions is known.}    & 0   & 120:24 & 5760 & ?   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withNa}  \\
    POPC:POPS (5:1)/Na$^+$  & Lipid14/17 \cite{dickson14,gould18} &2\todoi{Concentration to be checked after the amount of ions is known.}    & 0   & 120:24 & 5760 & ?   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withNa}  \\
    POPC:POPS (5:1)/Na$^+$  & Lipid14/17 \cite{dickson14,gould18} &3\todoi{Concentration to be checked after the amount of ions is known.}    & 0   & 120:24 & 5760 & ?   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withNa}  \\
    POPC:POPS (5:1)/Na$^+$  & Lipid14/17 \cite{dickson14,gould18} &4\todoi{Concentration to be checked after the amount of ions is known.}    & 0   & 120:24 & 5760 & ?   & 298  & 300 & 200 & \cite{POPCpopsLIPID17withNa}  \\

@@

    POPC:POPS (5:1)/Na$^+$ & Lipid14/17 \cite{dickson14,gould18} &0\todoi{Concentration to be checked after the amount of ions and water is known.}    & 0.03 & 60:12 & ?     & ?   & 298  & ?  & ?  & \cite{??} \todoi{Data to be delivered by Melcr}  \\
    POPC:POPS (5:1)/Na$^+$ & Lipid14/17 \cite{dickson14,gould18} &0\todoi{Concentration to be checked after the amount of ions and water is known.}    & 0.17 & 60:12 & ?     & ?   & 298  & ?  & ?  & \cite{??} \todoi{Data to be delivered by Melcr}  \\
    POPC:POPS (5:1)/Na$^+$ & Lipid14/17 \cite{dickson14,gould18} &0\todoi{Concentration to be checked after the amount of ions and water is known.}    & 0.36 & 60:12 & ?     & ?   & 298  & ?  & ?  & \cite{??} \todoi{Data to be delivered by Melcr}  \\

@@

    POPC:POPS (4:1)/Na$^+$  & Berger \cite{tieleman99,mukhopadhyay04}\todoi{Are these correct references?} &1.03    & 0   & 102:26 & 4290 & 80  & 310  & 200 & 50 & \cite{POPCpopsBERGERwith1000mMNa}  \\
    POPC:POPS (4:1)  & Berger \cite{tieleman99,mukhopadhyay04} &  0  & 0.12\footnote{Calculation of concetration complicated due the scaled ions. Concentration taken as reported in the delivered data.}   & 104:24 & 4306 & 24 & 310  & 300 & 100 & \cite{POPCpopsBERGERwith102mMCa}  \\
    POPC:POPS (4:1)  & Berger \cite{tieleman99,mukhopadhyay04} &  0  & 0.715\footnote{Calculation of concetration complicated due the scaled ions. Concentration taken as reported in the delivered data.}  & 104:24 & 4306 & 72 & 310  & 300 & 100 & \cite{POPCpopsBERGERwith715mMCa}  \\

@@

Two different definitions for the salt concentrations have been used when
electrometer concept is applied to study ion binding affinity.
The concentrations are reported either in water before solvating the lipids \cite{akutsu81,roux90,catte16}
or in bulk water after solvating the lipids \cite{altenbach84,melcr18}.
In this work, we use the former definition to be consistent with the reference
experimental data \cite{roux90}. The used definition has only a marginal effect
to the results is simulations with realistic ion binding affinity.

\todo{Should we discuss about hydration levels in experiments and simulations?}

@@

To evaluate counterion binding in different simulation models against experimental data~\cite{roux90},
we plot the headgroup order parameters measured from POPC:POPS 5:1 mixture
as a function of different monovalent ions added to the buffer (Fig. \ref{PSresponseTONaCl}). 
Experimental order parameter data for POPC headgroup in the mixture is available as a function
of LiCl and KCl concentrations, while POPS headgroup order parameters are measured also
as a function of NaCl. Lithium interacts more strongly with PS headgroups than other monovalent 
ions~\cite{hauser83,hauser85,roux86,mattai89,roux90}, as also observed for PC headgroups~\cite{cevc90}. 
This is evident also in the changes of PS headgroup order parameters, which decrease with the addition of lithium 
but increase with the addition of sodium or potassium (Fig. \ref{PSresponseTONaCl}). 
POPC headgroup order parameters exhibit a clear decrease as a function of LiCl concentration
but only modest changes as a function of KCl concentration, indicating singificant 
Li$^+$ binding but only weak Na$^+$ binding to the mixture when interpreted using the
electrometer concept~\cite{akutsu81,altenbach84,seelig87}. In simulations with the
Berger model, the headgroup order parameter response of POPC
to the added NaCl is similar to the experiments of LiCl,
indicating overestimated binding affinity of sodium, in line with the results for PC bilayers \cite{catte16}.
Indeed, the sodium density profile shows a significant binding peak in the 
Berger model (Fig. \ref{CIdensPSOCmixt}). Potassium binding in the MacRog simulation
is significantly weaker  (Fig. \ref{CIdensPSOCmixt}) and the headgroup order parameter 
changes are also in better agreement with simulations (Fig. \ref{PSresponseTONaCl}).
\todo{Discussion about Lipid17 to be written when we have the density profiles.}
All the tested models overestimate the changes of POPS headgroup order parameters as
a function of monovalent ions (Fig. \ref{PSresponseTONaCl}), suggesting that
model development is necessary to interpret the PS headgroup-ion interactions 
from MD simulations.
\begin{figure*}[]
  \centering
  \includegraphics[width=18.0cm]{../Figs/CHANGESwithMONVALENTwithPS.eps}
  \caption{\label{PSresponseTONaCl}
    Changes of the PC (left) and PS (right) headgroup order parameters as a function of
    added NaCl, KCl and LiCl from POPC:POPS (5:1) mixture at 298~K
    (except Berger simulations are (4:1) mixture at 310~K).
    The experimental data is from Ref. \citenum{roux90}.
    The values from counterion-only systems are set as a zero point of y-axis.
    To correctly illustrate the significant forking of the $\alpha$-carbon order parameter
    in PS headgroup (bottom, right), the y-axis is tranferred with the same value for both order parameters such that the lower order
    parameter value is at zero.
  }
  \todo{CHARMM36 results for this plot would be highly useful.}
\end{figure*}

### 2018-08-17 Manuscript/manuscriptPSsuppl.tex
Update text and figures

\pagebreak
\section{Calibration of PC headgroup order parameter response to the bound cations}\label{electrometerCALIBRATION}
Before using the headgroup order parameters to compare ion binding affinity between simulations
and experiments, the response of the order parameters to the bound charge in simulations should be quantified
against experiments~\cite{catte16,melcr18}. In our previous work~\cite{catte16},
the ratio $\Delta S_{\rm CH}^\beta / \Delta S_{\rm CH}^\alpha$ was in good agreement with the experiments~\cite{akutsu81}
in the Lipid14 model, but was underestimated by other models. In the more recent study~\cite{melcr18},
the headgroup order parameter responses were compared more carefully with the experiments of cationic
surfactants in POPC bilayer~\cite{scherer89}. The comparison reveals that the both headgroup
order parameters in the Lipid14 model are too sensitive to the bound charge,
while CHARMM36 gives better agreement for the $\alpha$ carbon (Fig. \ref{CHANGESwithDHMDMAB}).
Therefore, the headgroup order parameter response to the bound charge is actually more realistic
in CHARMM36 model than in the Lipid14, where both order parameters are equally too 
sensitive to the bound cations giving a good result for the $\Delta S_{\rm CH}^\beta / \Delta S_{\rm CH}^\alpha$ ratio.
The ratio was overestimated for the CHARMM36 model because the $\beta$-carbon order parameter
is relatively more sensitive than the $\alpha$-carbon order parameter. These results have to be taken
into account when analysing the ion binding affinities using headgroup order parameters in simulations.
However, we conclude that the discrepancies arising from the sentivity of lipid headgroup to
bound charge are typically smaller than then discrepancies arising from ion binding affinity.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGopsDHMDMAB.eps}
  \caption{\label{CHANGESwithDHMDMAB}
  Responses of headgroup order parameters to the fixed amount of cationic surfactants in
  POPC bilayer from simulations and experiments \cite{scherer89}.
  The simulation results for Lipid14 are directly from Ref.~\citenum{melcr18}.
  The CHARMM36 simulation data and details are available from Ref.~\citenum{CHARMM36cationicSURF}.
}
\end{figure}

>>>>

\pagebreak
\section{Calibration of PC headgroup order parameter response to the bound cations}\label{electrometerCALIBRATION}
Before using the headgroup order parameters to compare ion binding affinity between simulations
and experiments, the response of the order parameters to the bound charge in simulations should be quantified
against experiments~\cite{catte16,melcr18}. In our previous work~\cite{catte16},
the ratio $\Delta S_{\rm CH}^\beta / \Delta S_{\rm CH}^\alpha$ was in good agreement with the experiments~\cite{akutsu81}
in the Lipid14 model, but was underestimated by other models. In the more recent study~\cite{melcr18},
the headgroup order parameter responses were compared more carefully with the experiments of cationic
surfactants in POPC bilayer~\cite{scherer89}. The comparison reveals that the both headgroup
order parameters in the Lipid14 model are too sensitive to the bound charge,
while CHARMM36 gives better agreement for the $\alpha$ carbon (Fig. \ref{CHANGESwithDHMDMAB}).
Therefore, the headgroup order parameter response to the bound charge is actually more realistic
in CHARMM36 model than in the Lipid14. In the latter, both order parameters are equally too 
sensitive to the bound cations giving a good result for the $\Delta S_{\rm CH}^\beta / \Delta S_{\rm CH}^\alpha$ ratio.
The ratio was overestimated for the CHARMM36 model because the $\beta$-carbon order parameter
is relatively more sensitive than the $\alpha$-carbon order parameter. These results have to be taken
into account when analysing the ion binding affinities using headgroup order parameters in simulations.
However, we conclude that the discrepancies arising from the sentivity of lipid headgroup to
bound charge are typically smaller than then discrepancies arising from ion binding affinity.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGopsDHMDMAB.eps}
  \caption{\label{CHANGESwithDHMDMAB}
  Responses of headgroup order parameters to the fixed amount of cationic surfactants in
  POPC bilayer from simulations and experiments \cite{scherer89}.
  The simulation results for Lipid14 are directly from Ref.~\citenum{melcr18}.
  The CHARMM36 simulation data and details are available from Ref.~\citenum{CHARMM36cationicSURF}.
}
\end{figure}
